Essays in Computational Management Science

### 2018-07-18 Chapters/sdm2.tex
properties of the space

The space studied by Kanerva is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}. For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance. That is, $n$-sized bitstrings correspond to nodes, and edges exist between nodes $iff$ they flip a single bit.  Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community. A good survey is provided by \citet{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.

@@

>>>>

The space studied by Kanerva is also called the \emph{hypercube graph}, or $Q_n$, as in Figure \ref{hypercubes}. For a fixed $n \in \mathbb{Z}$, the graph $G = (V, E)$, in which $v \in V$ iff there is a bijective function $b: V\to \{0,1\}^n$ and $(v_i, v_j) \in E$ iff $H(b(v_i), b(v_j))=1$, where $H$ is the Hamming distance. That is, $n$-sized bitstrings correspond to nodes, and edges exist between nodes $iff$ they flip a single bit.  Though Kanerva has derived many combinatorial properties of the space, additional results have been found by the graph-theoretical community: it is a bipartite graph with chromatic number 2, and it is planar only if $n\le3$. The proofs, though elementary, illuminate some of its properties:  vertices connect bitstrings with an even number of 1's with bitstrings with an odd number of 1's, therefore (i) it is \emph{bipartite} and (ii) it \emph{may be colored with two colors}.  As for planarity, the case $n=3$ is the largest planar one.  Imagine a square inside a larger square, then connect the corresponding edges in both squares to obtain a planar drawing.  $Q_4$ or larger demands three such squares, with the smallest and largest connected but not passing through the middle square.

A good survey is provided by \citet{harary1988survey} --- and some interesting results may be found in \citep{foldes_characterization_1977, wagner_embedding_1990, laborde_another_1982, ruskey_combinatorial_2003}.

@@

A major difference is that in the 3-d world, there exist very far points, while in $Q_n$ the largest possible distance is $n$.  Another one is that the number of directions in $Q_n$ is enormous:  at \emph{every point} there are $n$ directions to choose from.
